---
---

## Exercises

===

### Exercise 1

Starting from {{ site.data.lesson.handouts[1] }}, modify the call to `renderText()` to
create an `output[['species_label']]` with the genus and species name. You can
find those data in the species table using the same "species_id". Hint: The
function `paste()` with argument `collapse = ' '` will convert a data frame row
to a text string.

[View solution](#solution-1)
{:.notes}

===

### Exercise 2

Modify the app completed in {{ site.data.lesson.handouts[2] }} to include a `tabsetPanel()`
nested within the main panel. Title the first tab in the tabset "Plot" and show
the current plot. Title the second tab in the tabset "Data" and show a
`dataTableOutput()`, which you can borrow from the app completed in {{
site.data.lesson.handouts[3] }}.

[View solution](#solution-2)
{:.notes}

===

### Exercise 3

Use the `img` builder function to add a logo, photo, or other image to your app
in {{ site.handoouts[2] }}. The help under `?img` states that HTML attributes
come from named arguments to `img`, and the ["img" HTML
tag](https://www.w3schools.com/tags/tag_img.asp) requires two attributes, and
you'll probably also want to set a valide `width`. Hint: Use the `www/images`
folder mentioned in the `addResourcesPath()` function at the bottom of {{
dite.handouts[2] }}.

[View solution](#solution-3)
{:.notes}

===

### Exercise 4

Notice the exact same code exists twice within the server function of the app
you completed in {{ site.data.lesson.handouts[3] }}: once for `renderPlot()` and once for
`renderDataTable`. The server has no way to identify an intermediate result, the
filtered data frame, which it could have reused. Replace `slider_months()` with
a new `selection_animals()` function that returns the entire `data.frame` needed
by both outputs. Bask in the knowledge of the CPU time saved, and congratulate
yourself for practicing DYR!

[View solution](#solution-4)
{:.notes}

===

## Solutions

===

### Solution 1

```{r, eval=FALSE, handout = 1}
# Packages
library(dplyr)
library(ggplot2)

# Data
species <- read.csv('data/species.csv',
  stringsAsFactors = FALSE)
animals <- read.csv('data/animals.csv',
  na.strings = '',
  stringsAsFactors = FALSE)

# User Interface
in1 <- selectInput(
  inputId = 'pick_species',
  label = 'Pick a species',
  choices = unique(species[['id']]))
out1 <- textOutput('species_label')
out2 <- plotOutput('species_plot')
tab <- tabPanel(
  title = 'Species',
  in1, out1, out2)
ui <- navbarPage(
  title = 'Portal Project',
  tab)

# Server
server <- function(input, output) {
  output[['species_label']] <- renderText({
    species %>%
      filter(id == input[['pick_species']]) %>%
      select(genus, species) %>%
      paste(collapse = ' ')
  })
  output[['species_plot']] <- renderPlot({
    df <- animals %>%
      filter(species_id == input[['pick_species']])
    ggplot(df, aes(year)) +
      geom_bar()
  })
}

# Create the Shiny App
shinyApp(ui = ui, server = server)
```

[Return](#exercise-1)
{:.notes}

===

### Solution 2

```{r, eval=FALSE, handout = 2}
# Packages
library(dplyr)
library(ggplot2)

# Data
species <- read.csv('data/species.csv',
  stringsAsFactors = FALSE)
animals <- read.csv('data/animals.csv',
  na.strings = '',
  stringsAsFactors = FALSE)

# User Interface
in1 <- selectInput(
  inputId = 'pick_species',
  label = 'Pick a species',
  choices = unique(species[['id']]))
side <- sidebarPanel('Options', in1)
out1 <- textOutput('species_label')
out2 <- tabPanel(
  title = 'Plot',
  plotOutput('species_plot'))
out3 <- tabPanel(
  title = 'Data',
  dataTableOutput('species_table'))                 
main <- mainPanel(out1, tabsetPanel(out2, out3))
tab <- tabPanel(
  title = 'Species',
  sidebarLayout(side, main))
ui <- navbarPage(
  title = 'Portal Project',
  tab)

# Server
server <- function(input, output) {
  output[['species_label']] <- renderText({
    species %>%
      filter(id == input[['pick_species']]) %>%
      select(genus, species) %>%
      paste(collapse = ' ')
  })
  output[['species_plot']] <- renderPlot({
    df <- animals %>%
      filter(species_id == input[['pick_species']])
    ggplot(df, aes(year)) +
      geom_bar()
  })
  output[['species_table']] <- renderDataTable({
    animals %>%
      filter(species_id == input[['pick_species']])
  })
}

# Create the Shiny App
addResourcePath('images', 'www/images')
shinyApp(ui = ui, server = server)
```

Notice the many features of the data table output. There are many options that
can be controlled within the render function such as pagination and default
length. See [here](http://shiny.rstudio.com/gallery/datatables-options.html) for
examples and how to extend this functionality using JavaScript.
{:.notes}

[Return](#exercise-2)
{:.notes}

===

### Solution 3

```{r, eval=FALSE, handout = 2}
# Packages
library(dplyr)
library(ggplot2)

# Data
species <- read.csv('data/species.csv',
  stringsAsFactors = FALSE)
animals <- read.csv('data/animals.csv',
  na.strings = '',
  stringsAsFactors = FALSE)

# User Interface
in1 <- selectInput(
  inputId = 'pick_species',
  label = 'Pick a species',
  choices = unique(species[['id']]))
img <- img(
  src = 'images/gkr2.jpg',
  alt = 'Portal Project research subject',
  width = '100%')
side <- sidebarPanel(img, 'Options', in1)
out1 <- textOutput('species_label')
out2 <- tabPanel(
  title = 'Plot',
  plotOutput('species_plot'))
out3 <- tabPanel(
  title = 'Data',
  dataTableOutput('species_table'))                 
main <- mainPanel(out1, tabsetPanel(out2, out3))
tab <- tabPanel(
  title = 'Species',
  sidebarLayout(side, main))
ui <- navbarPage(
  title = 'Portal Project',
  tab)

# Server
server <- function(input, output) {
  output[['species_label']] <- renderText({
    species %>%
      filter(id == input[['pick_species']]) %>%
      select(genus, species) %>%
      paste(collapse = ' ')
  })
  output[['species_plot']] <- renderPlot({
    df <- animals %>%
      filter(species_id == input[['pick_species']])
    ggplot(df, aes(year)) +
      geom_bar()
  })
  output[['species_table']] <- renderDataTable({
    animals %>%
      filter(species_id == input[['pick_species']])
  })
}

# Create the Shiny App
addResourcePath('images', 'www/images')
shinyApp(ui = ui, server = server)
```

[Return](#exercise-3)
{:.notes}

===

### Solution 4

```{r, eval = FALSE, handout = 3}
# Packages
library(dplyr)
library(ggplot2)

# Data
species <- read.csv('data/species.csv',
  stringsAsFactors = FALSE)
animals <- read.csv('data/animals.csv',
  na.strings = '',
  stringsAsFactors = FALSE)

# User Interface
in1 <- selectInput(
  inputId = 'pick_species',
  label = 'Pick a species',
  choices = unique(species[['id']]))
in2 <- sliderInput(
  inputId = 'slider_months',
  label = 'Month Range',
  min = 1, max = 12,
  value = c(1, 12))
side <- sidebarPanel('Options', in1, in2)									    
out1 <- textOutput('species_label')
out2 <- tabPanel(
  title = 'Plot',
  plotOutput('species_plot'))
out3 <- tabPanel(
  title = 'Data',
  dataTableOutput('species_table'))                 
main <- mainPanel(out1, tabsetPanel(out2, out3))
tab1 <- tabPanel(
  title = 'Species',
  sidebarLayout(side, main))
ui <- navbarPage(
  title = 'Portal Project',
  tab1)

# Server
server <- function(input, output) {
  selected_animals <- reactive({
    animals %>%
      filter(species_id == input[['pick_species']]) %>%
      filter(
        month >= input[['slider_months']][1],
        month <= input[['slider_months']][2])
  })
  output[['species_label']] <- renderText({
    species %>%
      filter(id == input[['pick_species']]) %>%
      select(genus, species) %>%
      paste(collapse = ' ')
  })
  output[['species_plot']] <- renderPlot({
    ggplot(selected_animals(), aes(year)) +
      geom_bar()
  })
  output[['species_table']] <- renderDataTable({
    selected_animals()
  })
}

# Create the Shiny App
shinyApp(ui = ui, server = server)
```

[Return](#exercise-4)
{:.notes}
